---
title: "EDA HW#7"
author: "Tonghao Zhang"
date: "2016/11/6"
output: pdf_document
---

```{r,echo=FALSE}
library(ggplot2)
x<-c(576,635,558,578,666,580,555,661,651,605,653,575,545,572,594)
y<-c(339,330,281,303,344,307,300,343,336,313,312,274,276,288,296)
data3<-data.frame(X=x,Y=y)
# ggplot(data3,aes(x=X,y=Y))+geom_point()
```


```{r,include=FALSE}
rrline1<-function(x,y){
n<-length(x);nmod3<-n%%3;
if(nmod3==0) 
  n3<-n/3;
if(nmod3==1) 
  n3<-(n-1)/3;
if(nmod3==2) 
  n3<-(n+1)/3;
x.order<-order(x)
medxL<-median(x[x.order][1:n3])
medxR<-median(rev(x[x.order])[1:n3])
medyL<-median(y[x.order][1:n3])
medyR<-median(rev(y[x.order])[1:n3])
slope1<-(medyR-medyL)/(medxR-medxL)
int1<-median(y-slope1*x)
newy<-y-slope1*x-int1
sumres<-sum(abs(newy))
return(list(a=int1,b=slope1,sumres=sumres,res=newy))
}

run.rrline<-function(xx,yy,iter=5){
out.coef<-matrix(0,iter,3)
ll<-(1:length(xx))[!is.na(xx)&!is.na(yy)]
n<-length(ll);x<-xx[ll];y<-yy[ll];newy<-y
for(i in 1:iter){
  rr<-rrline1(x,newy)
  out.coef[i,]<-c(rr$a,rr$b,rr$sumres)
  newy<-rr$res
  }
dimnames(out.coef)<-list(format(1:iter),c("a","b","|res|"))
aa<-sum(out.coef[, 1]);bb<-sum(out.coef[, 2]);cc<-sum(abs(y-aa-bb*x))
res<-yy-aa-bb*xx
out.coef<-rbind(out.coef, c(aa, bb, cc))
#print(round(out.coef,5))
return(list(a=aa, b=bb,res=res,coef=out.coef))
}

RRfit3<-run.rrline(data3$X,data3$Y)
# out.testjk<-matrix(0,15,3)
# pv.testjk<-matrix(0,15,3) 
# for(i in 1:15){
#   newx<-data3$X[-i]
#   newy<-data3$Y[-i]
#   rr1<-run.rrline(newx,newy)
#   out.testjk[i,]<-rr1$coef[6,1:3]
#   }
# tmp.coefs<-c(-47.91398,0.59140,152.3226)
# for(j in 1:3){
#   pv.testjk[,j]<-15*tmp.coefs[j]-14*out.testjk[,j]
#   }
# jk.se<-apply(pv.testjk,2,sd)/sqrt(15)
# jk.se

out.jk<-rep(0,15)
value.jk<-rep(0,15)
for(i in 1:15){
  newx<-data3$X[-i]
  newy<-data3$Y[-i]
  rho<-cor(x=newx,y=newy,method = "pearson")
  out.jk[i]<-0.5*log((1+rho)/(1-rho))
}
rho<-cor(x=data3$X,y=data3$Y,method = "pearson")
value.jk<-15*0.5*log((1+rho)/(1-rho))-14*out.jk
jk.est<-mean(value.jk)
jk.se<-sd(value.jk)/sqrt(15)
jk.est
jk.se

B<-2000
theta.bootstrap<-rep(0,B)
for(i in 1:B){
  ii<-sample(1:15,15,replace=TRUE)
  newx<-data3$X[ii]
  newy<-data3$Y[ii]
  rho<-cor(x=newx,y=newy,method = "pearson")
  theta.bootstrap[i]<-0.5*log((1+rho)/(1-rho))
}
bt.est<-mean(theta.bootstrap)
bt.se<-sd(theta.bootstrap)
bt.est
bt.se
critical<-qt(0.975,14)
jk.CI<-c(jk.est-critical*jk.se,jk.est+critical*jk.se)
bt.CI<-c(bt.est-critical*bt.se,bt.est+critical*bt.se)
jk.CI
bt.CI
critical<-qnorm(0.975,sd=1/(12))
fish.jk<-c(jk.est-critical,jk.est+critical)
fish.jk
fish.bt<-c(bt.est-critical,bt.est+critical)
fish.bt
```

##EX3

Jacknife estimate is 0.9170373, standard deviation is 0.4406764. CI: -0.02811946  1.86219412

Bootstrap estiamte is 1.125376, standard deviation is 0.3808777. CI: 0.307378 1.943373

Fisher's CI using Jacknife estimate as $\rho$ is 0.753707 1.080368.

Fisher's CI using Bootstrap estimate as $\rho$ is 0.9620454 1.2887060.


```{r,include=FALSE}
profl<-c(2.1,1.7,14.4,57.4,66.2,75.2,4.1)
farm<-c(6.8,8.1,14.8,62.4,81.7,94.0,10.2)
sales<-c(8.4,8.4,27.0,37.4,53.3,74.3,10.7)
crafts<-c(1.4,1.4,30.9,63.3,80.7,87.9,5.5)
oper<-c(14.6,12.0,36.5,65.5,79.7,93.3,18.1)
serv<-c(7.9,3.7,36.4,65.6,80.8,87.8,11.4)
labor<-c(4.8,4.5,31.4,59.8,82.4,80.5,6.1)
hearing<-cbind(profl,farm,sales,crafts,oper,serv,labor)
colnames(hearing)<-c("profl","farm","sales","crafts","oper","serv","labor")
row.names(hearing)<-c("500","1000","2000","3000","4000","6000","normal")

mpol<-medpolish(hearing)
res<-mpol$res
roweff<-mpol$row
coleff<-mpol$col
mat.roweff<-matrix(roweff,7,7)
mat.coleff<-matrix(rep(coleff,7),byrow=T,nrow=7)
Meff<-mpol$overall
NB<-200
bmat<-matrix(0,NB,15)
colnames(bmat)<-c("overall","500","1000","2000","3000","4000","6000","normal","profl","farm","sales","crafts","oper","serv","labor")
#dimnames(bmat)<-list(format(1:NB), c("Meff",paste("a",format(1:nrow(infmort)),sep=""), paste("b",format(1:ncol(infmort)),sep=""), "mad(res)"))
for(i in 1:NB){
  xx<-matrix(sample(res,49,replace=T),7)
  newdata<-xx+mat.roweff+mat.coleff+Meff
  new<-medpolish(newdata)
  bmat[i,]<-c(new$overall,new$row,new$col)
}


xbars<-function(x){
  c(mean(x,na.rm=T),sd(x,na.rm=T))
}

critical<-qt(0.975,6)
mat<-round(apply(bmat,2,xbars),3)
CI<-matrix(0,2,15)
colnames(CI)<-c("overall","500","1000","2000","3000","4000","6000","normal","profl","farm","sales","crafts","oper","serv","labor")
CI[1,]<-mat[1,]-critical*mat[2,]
CI[2,]<-mat[1,]+critical*mat[2,]
```

##EX4

CIs are:
```{r,echo=FALSE}
CI
effect<-c(roweff,coleff,Meff)
data4<-data.frame(effect=effect)
```

Notched boxplots of the row & column effects.

```{r,include=TRUE}
ggplot.data<-data.frame(fit=as.vector(t(bmat)))
ggplot.data<-cbind(ggplot.data,x=c("overall","500","1000","2000","3000","4000","6000","normal","profl","farm","sales","crafts","oper","serv","labor"))
ggplot(ggplot.data,aes(x=x,y=fit))+geom_boxplot(notch = TRUE)

```

From boxplots, the effects of 1000, 3000, 4000, 500, 6000, normal, oper, overall are significantly different from 0.


The correlation matrix is:

```{r,echo=FALSE}
# ggplot.data<-data.frame(fit=as.vector(t(bmat)))
# ggplot.data<-cbind(ggplot.data,x=c("overall","500","1000","2000","3000","4000","6000","normal","profl","farm","sales","crafts","oper","serv","labor"))
# ggplot(ggplot.data,aes(x=x,y=fit))+geom_boxplot(notch = TRUE)

cor<-cor(bmat)
cor
```

From the correlation matrix we can see there is strong relationship in pairs of overall and all frequency effects, pairs of frequencies. 